---
title: "ym2-detail"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ym2-detail}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(loglai)
```

This vignette shows the thought process behind making synthetic_ym2() into an efficient package command, and goes into detail about how to use the individual functions needed for synthetic_ym2().

# synthetic_ym2() components

`synthetic_ym2()` requires 2 different components to run, those being a particle filter and specifically formatted parameters, under the arguments `filter` and `pars`. Running the command on its own will return with an error!

Now the amount of work needed to make your own particle filter and parameters that fit within the requirements of `synthetic_ym2()` is quite daunting! Instead of making your own, it is far easier to use the particle filter and parameters supplied in this package.

# `synthetic_ym2_pf()`: The particle filter

The `synthetic_ym2_pf()` command will generate a particle filter for use in `synthetic_ym2()`. It does not require any arguments, so creating one is simple.

```{r}
filter.example <- synthetic_ym2_pf()
```

If you want variation in your particle filter, `synthetic_ym2_pf()` has an argument that will let you change the randomization seed. An evolution function is contained inside the particle filter which does NOT have a static output, so the default particle filter has a seed of `3141` to keep it consistent as well.

```{r}
filter.example2 <- synthetic_ym2_pf(seed = 10000)
```

`filter.example` and `filter.example2` will provide different results when used later.

# Csnippet functions contained inside of `synthetic_ym2_pf()`

There are some functions present inside of `synthetic_ym2_pf()` that are probably not practical for normal use.

`dmeas_mm()` and `synthetic_ym2_evo()` are both functions that create a singular Csnippet. The functions have no arguments, so they cannot be changed unless you modify them after using the command.

If you are curious, `dmeas_mm()` creates a measurement component for the particle filter, `while synthetic_ym2_evo()` defines the evolution function responsible for the particle filter's potential randomness.

# `GPP()`: Calculating GPP for the DALEC2 model

The `GPP()` function is more useful in general use than the prior 2 Csnippet functions, but is still situational, mainly used for the creation of the particle filter `ym2_pf()`. It calculates gross primary productivity based on the Aggregated Canopy Model, using a wide array of required arguments.

```{r}
GPP(lai = 2.641, c_eff = 20, maxt = 12.83247, mint = 6.867838, Ca = 400.95, lat = 46.23, yearday = 268, rad = 11.26568)
```

More details about these parameters can be found in the `GPP()` function's documentation.

# `synthetic_ym2_pars()`: Define default parameters for `synthetic_ym2()`

Similar to the particle filter, the information required to define your own parameters for the DALEC2 model calculations is rather difficult. `synthetic_ym2_pars()` defines default parameters fairly easily, but does not have much to tweak unlike `synthetic_ym2_pf()`'s seed argument.

```{r}
params.scaled <- synthetic_ym2_pars()
params.unscaled <- synthetic_ym2_pars(scaled = FALSE)
```

The `scaled` argument is a logical argument that will determine if the parameters calculated will be properly scaled to a [0,1]\^10 hypercube.

# `synthetic_ym2`: Generate a synthetic trajectory and marginal log-likelihood from the DALEC2 model

The simplest way of running `synthetic_ym2()` will be using the default particle filter and parameters.

```{r}
filter.example <- synthetic_ym2_pf()
params.scaled <- synthetic_ym2_pars()

dalec <- synthetic_ym2(filter = filter.example, pars = params.scaled)
```

As long as we provide a proper particle filter, you can provide any filter you want.

```{r}
filter.example2 <- synthetic_ym2_pf(seed = 10000)

dalec2 <- synthetic_ym2(filter = filter.example2, pars = params.scaled)
```

If the parameters provided are not scaled, you will have to specify that.

```{r}
params.unscaled <- synthetic_ym2_pars(scaled = FALSE)

dalec3 <- synthetic_ym2(filter = filter.example, pars = params.unscaled, scaled = FALSE)
```

Since the `scaled` argument is `TRUE` by default, a warning will display whenever completely default arguments are used to ensure the user knows what they're getting into. If you want to disable that, set the `verbose` argument to false.

```{r}
dalec <- synthetic_ym2(filter = filter.example, pars = params.scaled, verbose = FALSE)
```

You can also specify the amount of particles to use in the particle filter by adjusting the `n.part` argument. It will be set to 200 by default.

```{r}
dalec4 <- synthetic_ym2(filter = filter.example, pars = params.scaled, n.part = 300, verbose = FALSE)
```

# `synthetic_ym2()` output

The output of `synthetic_ym2()` will create a list containing the marginal log-likelihood and the trajectory. You can extract these from the objective function with the following keywords:

```{r}
dalec$logLik       # marginal log-likelihood
plot(dalec$traj)   # trajectory
```
